Sending a Balloon to the Edge of Space

### 2018-11-14 paper/paper.tex
almost done with paper

\usepackage{setspace}
\usepackage{gensymb}

\begin{document}
\maketitle
\begin{abstract}
This paper details 6thsense, a mission to send a weather balloon to to
edge of space in order to study Earth's atmosphere. The weather balloon
is outfitted with a payload that includes a camera, inside and outside
temperature sensors, a barometer, multiple GPS tracking devices, as well
as a host of individual experiments. The balloon most likely reached an
altitude of 20,000 meters, into the stratosphere and ozone layer. The
balloon was successfully recovered after approximately one and a half
hours of flight.
\end{abstract}

@@

\setcounter{secnumdepth}{2} \setcounter{tocdepth}{2} \tableofcontents
\newpage
\onehalfspacing

@@

There are four main layers to the atmosphere: the troposphere,
stratosphere, mesosphere, and thermosphere. The \textbf{troposphere} is
the section of the atmosphere from roughly 0km to 15km, and is the layer
in which we live. It includes 75\% of the mass of all gases in the
atmosphere, consisting of 78\% nitrogen gas, 21\% oxygen, and less than
1\% of argon, carbon dioxide, and other constituents. Because it is both
the lowest in altitude and most massive layer, the pressure in the
troposphere is the greatest: 100kPa. As you increase altitude in the
troposphere, temperature decreases to about -60C. The
\textbf{stratosphere} is the section from roughly 15km to 50km. The
ozone layer, or the region in the atmosphere which absorbs most UV rays,
is within the stratosphere at about 20km in altitude. The pressure in
the stratosphere is 10kPa. The \textbf{mesosphere} is the region from
50km to 80km. The pressure in the mesosphere is 0.1 kPa. Finally, the
\textbf{thermosphere} is the highest in altitude of the four layers,
spanning from 80km into outer space. Here, the pressure is only 0.01kPa,
as it is the least dense of the layers. The few air molecules in the
thermosphere are heated to extremely high temperatures by the Sun's
radiation. However, because there are so few molecules, we cannot
measure an appreciable difference in overall temperature, so the
effective temperature of the thermosphere is less than that of the
mesosphere.

The layers relevant to our experiment are the troposphere and
stratosphere, as our balloon launch was designed to reach a maximum
altitude of 30km (about the middle of the stratosphere but past the
ozone layer). Therefore, we had to prepare the payload for temperatures
as low as -60C and pressure as low as 10kPa (in the stratosphere). We
expect to encounter the lowest temperature in the troposphere, around an
altitude of 10km. As the balloon rises from sea level to the top of the
troposphere, we expect to see temperature decrease because the ground
absorbs most of the sun's radiation (as opposed to the gas particles).
However, as the balloon enters the stratosphere, we expect an increase
in temperature due to the ozone layer's blockage of UV rays. The UV rays
that the ozone layer absorbs causes the gas particles in the
stratosphere to heat up, so temperature increases with altitude. As the
balloon gains altitude, the density of air around it decreases, so the
pressure decreases.

@@

When infrared radiation enters the atmosphere, short-wave infrared
radiation (700 - 1000 nm) does not effectively interact with greenhouse
gasses. The photons instead pass through the atmosphere and reach the
surface of the earth. Here, they are absorbed by the surface of the
earth and get reemitted as a lower energy, longer-waved infrared photon.
Greenhouse gasses can now absorb these lower energy photons and reemit
them. This absorption and reemission warms the troposphere by increasing
the kinetic energy the atmosphere. Additionally, greenhouse gasses can
reemit photons back towards the surface of the earth, increasing the
temperature of the earth's surface. By impeding the tendency for
long-wave radiation to leave the atmosphere, the greenhouse effect
increase the temperature of the troposphere and surface.

The spectroscopy that explains the greenhouse effect concerns molecular
vibration. Infrared photons excite vibrations in molecules. Since the
vibrational frequency of atoms are quantized, only certain energies will
excite vibration in a molecule. Different molecules can exhibit
different vibrational modes. A molecule with more atoms has more
vibrational modes. The infrared selection rule states that a vibrational
mode is active when it is changing the dipole moment of the molecule. An
active vibrational mode is a molecular vibration that can be excited by
an infrared photon. Thus, infrared photons will only be absorbed when
that molecule's dipole moment is changing. Molecules such as nitrogen
and oxygen will never exhibit a change in their dipole moment due to
vibration because their only vibrational mode involves the singular
oscillation of the distance between two atoms. Thus, they cannot absorb
infrared radiation in order to partake in the greenhouse effect and are
not considered greenhouse gasses. R On the other hand, many atmospheric
molecules that have more than two atoms have more complex vibrational
modes. These active vibrational mode gasses, such as nitrogen dioxide
and water, are called greenhouse gasses.

As the greenhouse effect warms the planet, many factors accelerate the
temperature rise by creating positive feedback loops. Positive feedback
loops amplify any global increase in temperature. One positive feedback
loops is created when rising temperatures melt glaciers. These white and
reflective icy surfaces melt into dark oceans which absorb more
radiation and heat up faster. This explains the approaching transition
from a permanent to a seasonal ice cover in the Arctic Ocean. Thus
temperature increases accelerate.\footnote{\url{https://www.nature.com/articles/s41598-017-08467-z}}
Another positive feedback loops involves the melting of permafrost. A
quarter of the northern hemisphere is covered in permafrost that holds
concentrated methane and carbon dioxide. Slight growths in temperature
are enough to thaw the permafrost. When it melts, these greenhouse
gasses are released into the atmosphere. Thus, the greenhouse effect
amplifies and global temperature growth accelerates. \footnote{\url{http://science.sciencemag.org/content/340/6129/183}}
Finally, humidity is also a positive feedback loop. As global
temperatures rise, the atmosphere can hold more water vapor because less
of it gets condensed. Water vapor is a greenhouse gas itself. Thus any
increase in global temperatures will be amplified due to the water vapor
positive feedback loop.

The annual increase of atmospheric carbon dioxide concentration is
approximately 100 times greater in the past 60 years than it has been
during any previous natural increase.F Additionally, methane
concentrations have more than doubled since the start of the industrial
revolution.B Partly due to the increase of these gas concentration along
with many positive feedback loops, the planet's average surface
temperature has risen 0.9\degree C since the late 19th century.F With
rising temperatures comes the consequences of severe weather patterns,
flooding and other natural disasters. Collecting atmospheric data of
things such as temperature and gas concentrations informs our
understanding of the progression of the greenhouse effect. With this
information, we can begin to look at ways to impede rising global
temperatures.

The Panel on Climate Change has recognized that as global temperatures
rise, lightning strikes will occur more frequently. As our skyscrapers
get higher and multiply, lightning strike sourced Terrestrial Gamma ray
Flashes (TGF) become a concern.

@@

While there is an abundance of research done on upward directed TGFs
(those detected by the satellites in the 1950s), there is a lack of data
detecting downward directed TGFs. Currently, researchers led by David M.
Smith at UC Santa Cruz and other researchers around the world are using
a multitude of methods to detect downwards directed TGFs. Some methods
include sending detectors in cargo planes, on weather balloons and on
the ground in thunder prone areas. Since TGFs are so powerful, detectors
usually fail if they are near a TGF because of the sheer amount of
energy exerted. This field of research is relatively new and active.

\textbf{INSERT PRESSURE MODEL ANALYSIS HERE}

The exponential model for temperature vs altitude, where altitude is
less than 11km (or where altitude is within the troposphere) is:
T=T0(1-h/44329), where h=height in meters, and T0=temperature at sea
level. This relationship shows a negative relationship between height
and temperature. As altitude increases within the troposphere,
temperature decreases.

@@

@@

*shortly before launch it was discovered that the Geiger Counter was
non-operable and therefore it was taken out of the payload. Instead,
Simon Oros focused on \_\_\_\_.

\textbf{INCLUDE INDIVIDUAL MOTIVATIONS HERE}

@@

lorem ipsum

@@

lorem ipsum

@@

lorem ipsum

\section{Tracking The Payload}\label{tracking-the-payload}

>>>>

\usepackage{setspace}
\usepackage{gensymb}
\usepackage{pdfpages}

\begin{document}
\maketitle
\begin{abstract}
This paper details 6thsense, a mission to send a weather balloon to to
edge of space in order to study Earth's atmosphere. The weather balloon
is outfitted with a payload that includes a camera, inside and outside
temperature sensors, a barometer, multiple GPS tracking devices, as well
as a host of individual experiments. The balloon most likely reached an
altitude of 20,000 meters, into the stratosphere and ozone layer. The
balloon was successfully recovered after approximately one and one half
hours of flight.
\end{abstract}

@@

\setcounter{secnumdepth}{3} \setcounter{tocdepth}{3} \tableofcontents
\newpage
\onehalfspacing

@@

\subsection{Background}\label{background}

There are four main layers to the atmosphere: the troposphere,
stratosphere, mesosphere, and thermosphere. The \textbf{troposphere} is
the section of the atmosphere from roughly 0km to 15km, and is the layer
in which humans live. It includes 75\% of the mass of all gases in the
atmosphere, consisting of 78\% nitrogen gas, 21\% oxygen, and less than
1\% of argon, carbon dioxide, and other constituents. Because it is both
the lowest in altitude and most massive layer, the pressure in the
troposphere is the greatest: 100kPa. As altitude increases in the
troposphere, temperature decreases to about -60 \degree C. The**
stratosphere** is the section from roughly 15km to 50km. The ozone
layer, or the region in the atmosphere which absorbs most UV rays, is
within the stratosphere at about 20km in altitude. The pressure in the
stratosphere is 10kPa. The mesosphere is the region from 50km to 80km.
The pressure in the mesosphere is 0.1 kPa. Finally, the thermosphere is
the highest in altitude of the four layers, spanning from 80km into
outer space. Here, the pressure is only 0.01kPa, as it is the least
dense of the layers. The few air molecules in the thermosphere are
heated to extremely high temperatures by the Sun's radiation. However,
because there are so few molecules, an appreciable difference in overall
temperature cannot be measured, so the effective temperature of the
thermosphere is less than that of the mesosphere.

The layers relevant to the experiments conducted are the troposphere and
stratosphere, as the balloon was designed to reach a maximum altitude of
30km (about the middle of the stratosphere but past the ozone layer).
Therefore, the payload had to prepared for temperatures as low as -60
\degree C and pressure as low as 10kPa (in the stratosphere). The
payload is expected to encounter the lowest temperature in the
troposphere, around an altitude of 10km. As the balloon rises from sea
level to the top of the troposphere, temperature is expected to decrease
because the ground absorbs most of the sun's radiation (as opposed to
the gas particles). However, as the balloon enters the stratosphere, an
increase in temperature is expected due to the ozone layer's blockage of
UV rays. The UV rays that the ozone layer absorbs causes the gas
particles in the stratosphere to heat up, so temperature increases with
altitude. As the balloon gains altitude, the density of air around it
decreases, so the pressure decreases.

\subsection{Greenhouse Gases in the
Atmosphere}\label{greenhouse-gases-in-the-atmosphere}

@@

When infrared radiation enters the atmosphere, short-wave infrared
radiation (700 - 1000 nm) does not effectively interact with greenhouse
gasses. The photons instead pass through the atmosphere and reach the
surface of the earth. Here, they are absorbed by the surface of the
earth and get reemitted as a lower energy, longer-waved infrared photon.
Greenhouse gasses can now absorb these lower energy photons and reemit
them. This absorption and reemission warms the troposphere by increasing
the kinetic energy the atmosphere. Additionally, greenhouse gasses can
reemit photons back towards the surface of the earth, increasing the
temperature of the earth's surface. By impeding the tendency for
long-wave radiation to leave the atmosphere, the greenhouse effect
increase the temperature of the troposphere and surface. (Britannica,
2018)

The spectroscopy that explains the greenhouse effect concerns molecular
vibration. Infrared photons excite vibrations in molecules. Since the
vibrational frequency of atoms are quantized, only certain energies will
excite vibration in a molecule. Different molecules can exhibit
different vibrational modes. A molecule with more atoms has more
vibrational modes. The infrared selection rule states that a vibrational
mode is active when it is changing the dipole moment of the molecule. An
active vibrational mode is a molecular vibration that can be excited by
an infrared photon. Thus, infrared photons will only be absorbed when
that molecule's dipole moment is changing. (Pecsok, 1968) Molecules such
as nitrogen and oxygen will never exhibit a change in their dipole
moment due to vibration because their only vibrational mode involves the
singular oscillation of the distance between two atoms. Thus, they
cannot absorb infrared radiation in order to partake in the greenhouse
effect and are not considered greenhouse gasses. On the other hand, many
atmospheric molecules that have more than two atoms have more complex
vibrational modes. These active vibrational mode gasses, such as
nitrogen dioxide and water, are called greenhouse gasses. (UCAR, 2012)

As the greenhouse effect warms the planet, many factors accelerate the
temperature rise by creating positive feedback loops. Positive feedback
loops amplify any global increase in temperature. One positive feedback
loop is created when rising temperatures melt glaciers. These white and
reflective icy surfaces melt into dark oceans which absorb more
radiation and heat up faster. This explains the approaching transition
from a permanent to a seasonal ice cover in the Arctic Ocean. Thus
temperature increases accelerate. (Kashiwase, 2017)

Another positive feedback loop involves the melting of permafrost. A
quarter of the northern hemisphere is covered in permafrost that holds
concentrated methane and carbon dioxide. Slight growths in temperature
are enough to thaw the permafrost. When it melts, these greenhouse
gasses are released into the atmosphere. Thus, the greenhouse effect
amplifies and global temperature growth accelerates. (Vaks, 2013)

Finally, humidity is also a positive feedback loop. As global
temperatures rise, the atmosphere can hold more water vapor because less
of it gets condensed. Water vapor is a greenhouse gas itself. Thus any
increase in global temperatures will be amplified due to the water vapor
positive feedback loop.

The annual increase of atmospheric carbon dioxide concentration is
approximately 100 times greater in the past 60 years than it has been
during any previous natural increase (Lindsey, 2018). Additionally,
methane concentrations have more than doubled since the start of the
industrial revolution (Britannica, 2018). Partly due to the increase of
these gas concentration along with many positive feedback loops, the
planet's average surface temperature has risen 0.9\degree C since the
late 19th century (Lindsey). With rising temperatures comes the
consequences of severe weather patterns, flooding and other natural
disasters. Collecting atmospheric data of things such as temperature and
gas concentrations informs our understanding of the progression of the
greenhouse effect. With this information, methods can be researched and
developed to impede rising global temperatures.

The Panel on Climate Change has recognized that as global temperatures
rise, lightning strikes will occur more frequently. As skyscrapers get
higher and multiply, lightning strike sourced Terrestrial Gamma ray
Flashes (TGF) become a concern.

@@

While there is an abundance of research done on upward directed TGFs
(those detected by the satellites in the 1950s), there is a lack of data
detecting downward directed TGFs. Currently, researchers led by David M.
Smith at UC Santa Cruz and other researchers around the world are using
a multitude of methods to detect downwards directed TGFs. Some methods
include sending detectors in cargo planes, on weather balloons and on
the ground in thunder prone areas. Since TGFs are so powerful, detectors
usually fail if they are near a TGF because of the sheer amount of
energy exerted. This field of research is relatively new and active.
(Smith, 2015)

\subsection{Empirical and Exponential Models for Temperature and
Pressure}\label{empirical-and-exponential-models-for-temperature-and-pressure}

\begin{figure}[h]

{\centering \includegraphics{assets/nasa_model}

}

\caption{\label{fig:figs} NASA Empirical Models for Temperatur and Pressure}\label{fig:NASA_image}
\end{figure}

Figure 1 shows NASA's empirical models for pressure and temperature for
the troposphere and stratosphere. For a height of 0, the temperature is
that of sea level, or 15.06 \degree C. At a height of 11km, the
predicted temperature is -56.35 \degree C, which is consistent with
measurements recorded by the payload. As the balloon rises through the
lower stratosphere, a constant temperature of -56.46 \degree C is
expected. Finally, once the balloon rises beyond the ozone layer (around
25km altitude), temperatures are expected to rise again due to the ozone
layer's absorption of the sun's radiation. Refer to section 8 for an
overlay of NASA's empirical models over experimental data.

@@

\subsubsection{How a Pressure Sensor
Works}\label{how-a-pressure-sensor-works}

Most pressure sensors, including the Vernier gas pressure sensor that is
connected to the main Arduino mega board and used on the payload,
measure gas pressure by using a flexible diaphragm, where one side is a
vacuum and the other side is open to the surrounding atmosphere, and
producing a voltage that varies linearly with absolute pressure. A
typical capacitive humidity sensor measures relative humidity by placing
a sheet of metal oxide between two electrodes and measuring the change
in capacitance between the electrodes, which is directly proportional to
the change in air moisture. And lastly, the temperature sensor used in
the following experiments, specifically a thermistor, measures
temperature through nonlinear changes in resistance, which decreases as
temperature increases.

@@

*shortly before launch it was discovered that the Geiger Counter was
non-operable and therefore it was taken out of the payload. Instead,
Simon Oros focused on ground-level radiation measurements.

\subsubsection{Alec Vercruysse:
Spectrometer}\label{alec-vercruysse-spectrometer}

Identifying the gaseous composition of the atmosphere is of great
importance in order to be able to properly analyze the different layers
and their properties. Specifically, identifying the composition of the
ozone layer is extremely important in order to determine its health.
Actively monitoring the ozone layer is crucial, since it is responsible
for shielding life on earth from harmful Ultraviolet light. Furthermore,
analyzing the gaseous composition of the troposphere is important to
understanding and measuring the greenhouse effect (see section 1). The
more greenhouse gasses in the atmosphere, the larger the greenhouse
effect. This leads to more heat trapped in the atmosphere, and
accelerated global warming.

A spectrometer was placed at a 45 \degree angle relative to the top
acrylic plate in order to perform spectral analysis of the atmosphere,
with a view unobstructed by the balloon or payload.

\subsubsection{John Kim: Methane Sensor}\label{john-kim-methane-sensor}

Greenhouse gases are gases that trap heat that come from both the human
activity and the Sun inside the Earth's atmosphere, resulting in global
warming and climate change when overlooked. While many identify
greenhouse gases with carbon dioxide, methane contributes to over 10\%
of the gaseous composition of the atmosphere and is emitted from coal
mining and organic waste. Monitoring methane levels is crucial to not
only furthering the understanding of how the Earth's atmospheric
composition changes over time, but also protecting the
environment---methane is much more efficient than carbon dioxide at
absorbing heat and more potent. Scientists have even written letters to
the United Nations' climate change summit to require nations to curb
methane emissions more quickly.

A methane sensor was placed inside the insulated payload because it was
found to be fairly sensitive to temperature when compared to other
components that were placed outside the styrofoam box. In addition,
there would not have been any significant difference in methane levels
inside and outside the payload because the non-restrictive airflow
inside the box would have allowed methane gas to pass through relatively
unobstructed. A typical methane sensor measures methane gas by detecting
the changes in resistance when the gas comes into contact with a
catalytic surface and heats the wiring resistance.

\subsubsection{Caroline Pang: UVB
Sensor}\label{caroline-pang-uvb-sensor}

Ultraviolet light is a harmful form of radiation which enters our
atmosphere. There are three classifications of ultraviolet radiation,
UVA, UVB, and UVC. The ozone layer is responsible for absorbing most of
the harmful UVC radiation making it an essential part of the atmosphere.
However, it only partially absorbs UVB light which is responsible for
causing sunburns and in extreme cases, skin cancer. Measuring the
intensity of UVB light as a function of altitude and observing how it
changes when entering the ozone layer can demonstrate how effectively
the ozone layer absorbs UVB light.

Therefore, a UVB light sensor was placed vertically inside the payload
with the top half of the rod outside of the acrylic top to measure the
intensity of UVB light.

\subsubsection{Vivian Liu: Light Intensity and Solar
Power}\label{vivian-liu-light-intensity-and-solar-power}

Light intensity, specifically for the purposes of solar energy, is an
important part of the initiative for cleaner energy. It is also
interesting to learn about how light intensity increases or decreases as
it passes through the ozone layer.

To test light intensity, a photoresistor was placed on a board, and the
resistances of the photoresistor attached in series to a 1.2k Ohm
resistor were recorded as a function of altitude. From the resistance of
the photoresistor, lux, or the units of light intensity, can be
calculated. For solar power, the arduino read the voltage of a solar
panel that was placed outside of the shadow of the balloon. From this
voltage, power output of the solar panel can be calculated. The purpose
of the experiments on solar properties is to see if there is a
correlation between altitude and solar output.

@@

\subsection{Payload Design}\label{payload-design}

\subsubsection{Structure}\label{structure}

The payload design consists of a styrofoam box and an acrylic rectangle
as a lid. Styrofoam was used because it is both light and insulating.
The acrylic lid was heavy, but it allowed the interior of the payload to
be seen at all times. Most importantly, it provided for firm mounting
points for the arduino, all circuit boards, and the sticky GoPro mount.
The lid has two rectangles cut out in the middle so that wires and
sensors can go from the outside to the inside of the payload. The lid
has two holes along its width to accommodate two vertical wooden dowels
that fasten the lid to the styrofoam box. A hole was drilled through the
wall of the styrofoam box so that the dowel can be inserted and secured
using epoxy glue. The wooden dowels are then doubly secured by fastening
it to the box using cotter pins and washers. The payload can be
connected to the balloon using the two wooden dowels that run
horizontally through the top of the payload. Strings are tied to the
ends of the wooden dowels, and secured using a combination of epoxy glue
and cotter pins/washers. Finally, the cutdown mechanism was secured by
wrapping wire around the string connecting the payload to the carabiner.

\subsubsection{Organization of Parts}\label{organization-of-parts}

Most of the important electronics were attached to the acrylic lid of
the payload. The arduino mega and protoboard containing individual
experiments are screwed onto the interior of the acrylic lid using nylon
spacers and screws. These components are on the interior of the payload
for protection. The BME sensor, protoboard (containing sensors), GoPro
mount, and transmitter are screwed onto the exterior of the acrylic lid.
The spectrometer is outside the payload to give it access to the sky,
and the transmitter is also on the exterior to provide better
line-of-sight. These essential electronics are secured onto the lid so
as to reduce movement within the payload. Other plastic items, such as
the UV sensor, gas pressure sensor, and fox-hunt transmitter, were
placed within the styrofoam box. Finally, the power source, four 9V
batteries, and the picoAPRS were placed at the bottom of the payload
underneath a layer of fiberglass insulation. This was to help preserve
thermal energy inside the payload so that the batteries could continue
working even in very cold conditions, including those of the PicoAPRS.

Near the bottom of the payload, there are two 3' carbon rods that jut
out of the payload, and a wooden box at the end of the rods. This is so
that the solar panel can be secured on top of the wood box, and the
distance from the balloon to the end of the rods can ensure that the
solar panel is not obscured by the shadow of the balloon.

\paragraph{Table 1: Weights of Items in
Payload}\label{table-1-weights-of-items-in-payload}

\begin{longtable}[]{@{}ll@{}}
\toprule
\begin{minipage}[b]{0.72\columnwidth}\raggedright\strut
Item\strut
\end{minipage} & \begin{minipage}[b]{0.22\columnwidth}\raggedright\strut
Weight (g)\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.72\columnwidth}\raggedright\strut
Cutdown mechanism (protoboard + 6 capacitors + wiring)\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright\strut
64\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.72\columnwidth}\raggedright\strut
PicoAPRS\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright\strut
59\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.72\columnwidth}\raggedright\strut
Fox-Hunt Transmitter\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright\strut
77\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.72\columnwidth}\raggedright\strut
4 batteries (40g/battery)\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright\strut
160\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.72\columnwidth}\raggedright\strut
\textbf{Payload cover} (arduino mega, arduino uno, two protoboards,
transmitter, pressure sensor, GoPro mount, ME, UV sensor, acrylic cover,
wiring)\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright\strut
604\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.72\columnwidth}\raggedright\strut
\textbf{Payload body} (styrofoam, two 3' carbon rods, solar panel, small
wood box, four 10'' wood dowels)\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright\strut
175\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.72\columnwidth}\raggedright\strut
Fiberglass insulation\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright\strut
30\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.72\columnwidth}\raggedright\strut
GoPro\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright\strut
200\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.72\columnwidth}\raggedright\strut
\textbf{Total}\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright\strut
1.369 kg\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\subsection{The Tethered Launch Reel-in
System}\label{the-tethered-launch-reel-in-system}

\subsubsection{Structure}\label{structure-1}

The structure of the reel in system consists of a sturdy, slightly
elevated base and two sides to hold the axle. The base is made from a
square frame with one additional plank for the motor to be mounted on.
The axle holders are each made from one vertical 2x2 post with a hole
through the top to fit a bearing, and two slanted supports to form a
triangular shape. In addition to this, there are two thin wooden pieces
screwed on the outside of each post to prevent the axle from popping out
of the frame. All of the pieces are secured with construction screws.

\subsubsection{Gears and Axle}\label{gears-and-axle}

The reel in system has two gears, a 72 tooth gear on the axle, and a 10
tooth gear mounted on the motor. Therefore, there is a theoretical
mechanical advantage of 7.2. However, the measured mechanical advantage
was around 5.9. This mechanical advantage is necessary so that the motor
can provide enough torque to reel in the balloon and so that the axle
does not spin too quickly.

\[MA = \frac{f_{out}}{f_{in}} = \frac{d_{in}}{d_{out}} = \frac{72}{10}\]

Securing the larger gear to the axle is one of the most important parts
of the reel in system. Therefore, there are two mechanisms in place to
keep the gear in place. First there are two small disks which fit snugly
onto the axle on each side of the gear which are screwed together,
somewhat clamping the gear to the axle. However, the disks are still
able to rotate and slide along the axle when force is applied. Thus, the
addition of a more secure mechanism was implemented by attaching a
wooden block to the axle, drilling a hole through the block and axle,
and then securing them together by a screw. The two small disks were
then attached to the block by screws.

In addition to this, the axle includes two larger wooden disks spaced
apart to set the boundaries for where the string is wrapped. One failure
point in the reel in system was that the string was wrapped unevenly
around the axle, causing it to tangle multiple times during the tethered
launch. In the future, the string should be moved side to side more
often when being wrapped.

\subsubsection{The Tensioning System}\label{the-tensioning-system}

Maintaining tension in the chain is essential for the reel in system to
work effectively because increases the contact between the chain and the
gears, and it protects against the chain falling off the gears at high
speeds. To implement a tensioning system, a thin wooden bar was first
installed across axle holders to mount the system across from the chain.
The system is comprised of an S-shaped piece of wood with a hole going
horizontally through the top ends where an axle was placed to hold a
sprocket. The bottom end of the wood was secured to the frame with a
piece of string acting as a pivot point. To apply a downward force on
the piece of wood, a compression spring was placed between the wood and
the crossbar and held in place by screws. In operation, the spring
pushed the gear onto the chain, creating tension in the chain.

\subsection{Fox-Hunting Yagi Antenna}\label{fox-hunting-yagi-antenna}

The 3 element yagi antenna used for Fox-Hunting is built from PVC
piping, cross connectors, a tape measure, hose clamps, a piece of 14 AWG
wire, and a coax cable.

All three elements, the were each measured and cut to the appropriate
length from the tape measure so that the antenna could receive the
147.065mHz signal from the Fox-Hunting transmitter. The reflector was
cut to 41.375 in, the driven element was comprised of two 17.75 in
segments, and the director was 35.125 in long. Additionally, the spacing
between each element was carefully measured when cutting the PVC pipes
to the correct length, with 12.5 in between the director and driven
element, and 8 in between the driven element and the reflector. However,
even though our Fox-Hunting was successful, the lengths and distances of
each element could have been fine-tuned to increase the sensitivity of
the antenna to the frequency of our Fox-Hunting transmitter. See section
6.5.

A coax cable had one end separated into the grounded shield and signal
conductor, each which was attached to an opposite end of the driven
element. A wire was also connected to the driven element to complete the
dipole. The other end of the coax cable included a BNC connector that
was attached to a radio receiver.

All physical CAD drawings are included in the appendix.

@@

\begin{figure}[h]

{\centering \includegraphics{assets/cutdown}

}

\caption{\label{fig:figs} Cutdown mechanism circuit design}\label{fig:cutdown_diagram}
\end{figure}

\subsection{Cutdown Method}\label{cutdown-method}

The cutdown mechanism functions by running electricity through nichrome
wire in order to burn the string that attaches the payload to the
balloon. Two 5 cm lengths of 26g nichrome wire were used. 5 loops of the
double stranded nichrome wire were wrapped around the the string that is
to be cut. \(R_{2}\) represents the nichrome wire.

\subsection{Charge Needed}\label{charge-needed}

A stand with a 1.4 kg weight attached (the expected maximum weight of
the payload) was used as a testing apparatus. First, a power supply was
used to test the cutdown mechanism. The apparatus was placed in a
-47\degree C freezer for 30 minutes before testing. We chose this
temperature because it is colder than the expected minimum temperature
that the payload will experience. After 30 minutes, we observed that it
took 6.6A for 0.71s in order to cut the string. We calculated the amount
of charge needed to cut the string.

\[I=\frac{\Delta Q}{\Delta t}\]

The amount of charge needed is 4.7C.

\subsection{Capacitor Bank}\label{capacitor-bank}

While this is the minimum amount of charge required for the cutdown to
work, significantly more charge was stored in order to combat leaky
capacitors and cold temperatures. Our capacitor bank is composed of
supercapacitors (\(C_{1}-C_{6}\)). Each capacitor stores Three
capacitors were wired up in series. This increased the voltage of the
capacitor bank threefold. We chose to increase the voltage in order to
combat the cold conditions of space that are not favorable to the
cutdown method of heating. Electrolytic capacitors lose voltage in cold
conditions. The max voltage of the capacitor bank is calculated below.

\[V_{T} = 3V_{C}\]

The voltage of the capacitor bank is 8.1V.

Putting capacitors in series reduces their capacitance. Thus, we added
three capacitors in parallel to the original three capacitors in order
to increase the capacitance. Thus, more charge could be stored. The
capacitance of the capacitor bank (C1-C6) is calculated below.

\[C_{T} = \frac{1}{\frac{1}{C_{1}} + \frac{1}{C_{3}} + \frac{1}{C_{5}}}+\frac{1}{\frac{1}{C_{2}} + \frac{1}{C_{4}} + \frac{1}{C_{6}}}\]
The capacitance of the capacitor bank is 16.7C. The total charge stored
in this capacitor bank if it were charged at 8.1V is calculated below.

\[V=\frac{Q}{C}\]

The total charge stored is 135C. This is significantly more than the
charge needed for the cutdown to work (4.7C). Although excessive, this
ensures that the cutdown will have enough charge to function.

\subsection{Transistor Relay System}\label{transistor-relay-system}

The cutdown circuit has two power sources. VCC is the constant voltage
source supplied by the arduino. V--\textgreater{} is the cutdown signal
outputted from the arduino. Both voltage sources are 5V. Q1 is a NPN
transistor. NPN transistors work by letting current flow from the
collector to the emitter when there is a voltage applied at the base. We
used a TIP31C NPN transistor. The TIP31C requires between 1.8V and 5V of
current at the base to go on. We used a resistor between the arduino and
transistor in order to draw enough current for the base of the
transistor. When the transistor turns on, the arduinos constant 5 volt
source can send current through the switching section of a relay (K1).
When the relay is turned on, the capacitor bank (C1 - C6) can discharge
through the nichrome wire (R2).

There are two reasons for using of both a transistor and a relay. First,
transistors tend to burn out more readily when their maximum collector
current is exceeded than when a relay's maximum carrying current is
exceeded. Both the relay we used and the transistor have a maximum
current allowance of 3A. The discharging of the capacitor bank releases
a very high current. The discharge section of the circuit would benefit
from running through a component that could handle more current before
breaking. This is why a relay was used for the cutdown discharge and not
a transistor. Additionally, the relay needs to collect more current to
go on than the transistor does. Arduino signals cannot supply the needed
current to power the magnetic coils of the relay. Thus, a transistor was
used.

\subsection{Circuit Walkthrough and Physical
Design}\label{circuit-walkthrough-and-physical-design}

First, the V--\textgreater{} sends a signal that goes over R1 and into
the base of Q1. Q1 is now on and current can flow from Vcc, through the
K1 collector, through the collector and emitter of Q1, and to GND. The
relay is now on. C1-6 can discharge over R2. R2 becomes hot and cuts the
circuit.

The circuit was soldered on a lightweight protoboard. Electrical tape
was used to wrap the entire mechanism. This way, the protoboard would
not come in contact with other components in the payload and cause a
short circuit. Refer to appendix for diagrams of the cutdown mechanism
design and more physical specifications.

\subsection{Cutdown Testing}\label{cutdown-testing}

Many tests were conducted where power supplies simulated the voltage
outputs of an arduino. Over five tests conducted in this manner were
shown to cut 1.4 kg weights down at room temperature. The cutdown was
left at 38\degree C for 1 hour and 9 minutes. The cutdown still worked
in these conditions.

Next, the cutdown was tested during the tethered launch. A helium
balloon was attached at the top of the payload. The cutdown went off
while the payload was still on the ground. Pressure sensor fluctuations
were so severe that they managed to drop below the specified cutdown
pressure. To correct for this, the arduino now tests for the average of
10 readings dropping below the cutdown pressure and not single
measurements. Thus, fluctuations will be averaged out and the pressure
sensor will not incorrectly trigger the cutdown.

The cutdown was tested on a seperate arduino in a vacuum chamber. The
code used for this test follows the averaging method described above and
can be found in the appendix. The cutdown successfully burnt through a
string to release a 1.4 kg weight when the vacuum chamber dropped below
1.08 kPa. 1.08 kPa is the cutdown pressure specified in the code. This
value was chosen because it was the pressure at 100,000 ft.

\subsection{Final Code}\label{final-code}

The pressure testing cutdown code was implement into the main arduino's
code. The cutdown trigger relies on more than just the pressure in the
final code. As safety mechanisms to ensure the cutdown doesn't go off
too early, the code tests for temperature and time also. The balloon is
expected to take longer than 1.5 hours before reaching 100,000 ft.
Therefore, the cutdown will not trigger until 1.5 hours have passed.
Additionally, the temperature of the atmosphere is less than
-10\degree C at 100,000 ft. Thus, the code will not let the cutdown
trigger unless the outside temperature readings are less than
-10\degree C. The code ensures that the pressure average reading, the
temperature reading and the time all fall in the cutdown parameters for
10 readings before triggering the cutdown. This prevents from faulty
cutdown triggers due to sensor fluctuations.

There are limitations to the code that prevents the cutdown from
triggering too early. If the timer or pressure and temperature sensors
fail, the cutdown might never go off. While this error was not
encountered because the balloon popped too early (see section 8), this
was a possibility.

@@

\subsection{Background}\label{background-1}
